Parameter estimation

### 2023-05-07 SFE-article/Sections/Results.tex
07.05

    \begin{itemize}
        \item Partition coefficient:\qquad\quad\qquad$k_m$
        \item Internal Diffusion coefficient: \quad$D_i$
        \item Axial Diffusion coefficient: \qquad$D_e^M$
        \item Standard Deviation: \qquad\qquad\quad$\sigma$
    \end{itemize}

    Moreover, the initial state estimation was performed together with parameter estimation. The concentration of solute in the solid phase is assumed to be constant and uniformly distributed. On the other hand, the solute concentration in the fluid phase should not follow the same assumption as the solid phase. During the preparation period the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed) along the extractor. As a result, the distribution of solutes mass in the fluid phase is assumed to not be uniform. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in appendix \ref{CH: IC_BC}. The linear distribution can be defined if the total mass of solute $m_{total}$ and partition coefficient $\tau$ are know.

@@

	To ensure that parameters found by the optimizer do not reach unrealistic values, an additional set of inequality constraints is introduced. The lower and upper bounds for each parameter are given in Table \ref{tab:Constraints}. The same Table contains the initial guesses for the optimizer. To ensure that the solution is a global solution multiple initial points have been tested.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lcccccc }
			\hline 
			Parameter		&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Lower bound		&0	  	& 0 	  	& 0 		& 76 		 	& 0 	   	& 0 \\ 
			Upper bound		&1 		& $+\infty$ & $+\infty$	& 100 			& 1 		& $+\infty$ \\ 
			Initial guess	&1	  	& 3 	  	& 1 		& 77 		 	& 0.65 	   	& 0.1 \\  \hline
		\end{tabular} }
	\caption{Constraints of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

@@

	The figure \ref{fig: estimation_results} show yield curves for each of experiments. The blue dots indicate the data point obtained as a measurement from the laboratory experiments. The black curve represents the yield curve obtained from the initial guess of the parameters. The blue curve correspond to yield curve obtained as a solution of the optimization problem. The curves on the left hand-side of the Figure \ref{fig: estimation_results}, are the cumulative yield curves, which represents the cumulative amount of the oil collected during each experiment. The Figures on the right hand-side, represent the derivative of the cumulative yield curves. As explained in the chapter \ref{CH: Parameter_estimation}, the derivatives of the cumulative measurements are independent on the previous measurement and should be used for the parameter estimation.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautobooktabular{Figures/Results_estimation/estimation.csv} }
		\caption{Parameter estimation results}
		\label{tab: Estimation_results}
	\end{table}

	The results of the parameter estimation can be found in Table \ref{tab: Estimation_results}. As can be noticed, the optimizer found that the partition factor should be as high as possible and moved values of $k_m$ for each experiment to unity, which was chosen as the upper bound.

	Values of the internal diffusion coefficients are distinguished for each experiment, with an increasing trend with respect to the fluid density. The order of magnitude of $D_i$ obtained from the optimization is similar to values found by other researchers. \citet{Reverchon1996} performed the parameter estimation for the extraction process of sage oil from seeds and reported $D_i \approx 6e-13$.

@@

	As the total amount of the oil present in the system is unknown, and as such it has been obtained from the optimizer. Given the dataset and the process model, the optimizer found that all most all the oil was extracted out of the seeds and that the best fit is obtained if the total mass of the oil reach a lower bound equal to 76 g.

	As explained above, and in Appendix \ref{CH: IC_BC}, the linear distribution of can be fully determined if the parameter $\tau$ is known. The split ratio for experiments conducted at $40 ^\circ C$ / 300 bar and $50 ^\circ C$ / 300 bar are similar to each other and close to 0.64. The experiments performed at $40 ^\circ C$ / 200 bar and $50 ^\circ C$ / 200 bar have $\tau$ equals to $0.69$ and $0.74$, respectively.

	The noise present in each dataset is quantified by parameter $\sigma$. It can be observed dataset obtained at $40 ^\circ C$ / 200 bar, $50 ^\circ C$ / 200 bar and $40 ^\circ C$ / 300 bar have similar value of $\sigma\approx0.34$. The visual investigation of the dataset obtained at $50 ^\circ C$ / 300 bar allow to explain higher value of $\sigma$ for this dataset. This dataset have higher deviation of the data points from the simulated curve.

		\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P200.pdf}
			\caption{Experiment at $40^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P200.pdf}
			\caption{Experiment at $50^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P300_org.pdf}
			\caption{Experiment at $40^\circ C$ and $300$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P300_org.pdf}
			\caption{Experiment at $50^\circ C$ and $300$ bar}
		\end{subfigure}
		\caption{Results of parameter fitting, with estimation of the initial state}
		\label{fig: estimation_results}
	\end{figure}

	The results obtained from the parameter estimation are used to develop a generalized process model. This model expresses the parameters $D_i$ and $D_e^M$ as functions of the operating conditions, such as temperature (T), pressure (P), and flow rate (F). Since the experiments were performed under varying temperature and pressure, the influence of these variables is taken into account. To simplify the problem, the correlations are written as functions of fluid density rather than temperature and pressure.

	While the experimental procedure and device can affect $\tau$ and $\sigma$, these parameters are not further considered. They are used only to define the initial conditions and are not included in the process model. Additionally, since the optimizer found that $k_m$ and $m_{total}$ have the same value, they are assumed to be constants and not considered further.

	The other two parameters ($D_i$ and $D_e^M$) are subject to regression to obtain correlations in terms of density. Figure \ref{fig:Regression} shows the first- and second-order polynomials used to fit $D_i$ and $D_e^M$ as a function of fluid density. Interestingly, these two parameters exhibit opposite behavior.

	\begin{figure}[!h]
	\centering
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 10.5cm 11cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
		\caption{First order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
		\label{fig:Regression_1}
	\end{subfigure}
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 10.5cm 11cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
		\caption{Second order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
		\label{fig:Regression_2}
	\end{subfigure}
	\caption{Results of parameter fitting, with estimation of the initial state} 
	\label{fig:Regression}
	\end{figure}

@@

	The increase in the density of the supercritical fluid can lead to an increase in the solubility of the solute and, consequently, an increase in the concentration gradient. This increased concentration gradient drives the solute to diffuse more rapidly through the matrix of the solid particles, resulting in an increase in the internal mass transfer coefficient. However, the effect of increased solubility on the internal mass transfer coefficient is greater than the resistance coming from higher density, resulting in an overall increase in the internal mass transfer coefficient with density.

@@

	Figure \ref{fig:Regression_1} show the result of fitting a second-order polynomial in density space. It can be observed that the fit obtained from quadratic functions seems to be represent both dataset better.

>>>>

    \begin{itemize}
        \item Partition coefficient:\qquad\quad\qquad$k_m$
        \item Internal Diffusion coefficient: \quad$D_i$
        \item Axial Diffusion coefficient: \qquad$D_e^M$
        \item Saturation concentration: \quad\qquad$C_{sat}$
        \item Standard Deviation: \qquad\qquad\quad$\sigma$
    \end{itemize}

    Moreover, the initial state estimation was performed together with parameter estimation. The concentration of solute in the solid phase is assumed to be constant and uniformly distributed. On the other hand, the solute concentration in the fluid phase should not follow the same assumption as the solid phase. During the preparation period the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved along the extractor (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed). As a result, the distribution of solute mass in the fluid phase is assumed to not be uniform. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in appendix \ref{CH: IC_BC}. The linear distribution can be defined if the total mass of solute $m_{total}$ and partition coefficient $\tau$ are know.

@@

	To ensure that parameters found by the optimizer do not reach unrealistic values, an additional set of inequality constraints is introduced. The lower and upper bounds for each parameter are given in Table \ref{tab:Constraints}. To ensure that the solution is a global solution multiple initial points have been tested but presented results come from the initial guess given also in Table \ref{tab:Constraints}.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lccccccc }
			\hline 
			Parameter		&$k_m$[-] 	& $D_i$[m/s$^2$] 	& $D_e^M$[m/s$^2$]	& $C_{sat}$ [kg/m$^3$] 	& $m_{total}$[g]	& $\tau$[-] 	& $\sigma$[-] \\  \hline
			Lower bound		&0	  		& 0 	  			& 0 				&	0					& 80 		 		& 0 	   		& 0 \\ 
			Upper bound		&1 			& $+\infty$ 		& $+\infty$			&	$+\infty$			& 150 				& 1 			& $+\infty$ \\ 
			Initial guess	&1	  		& 3 	  			& 1 				&	3					& 80 		 		& 0.65 	   		& 0.1 \\  \hline
		\end{tabular} }
	\caption{Constraints of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

@@

	Figure \ref{fig: estimation_results} show yield curves for each of experiments. The blue dots indicate the data point obtained as a measurement from the laboratory experiments. The black curve represents the yield curve obtained from the initial guess of the parameters. The blue curve correspond to yield curve obtained as a solution of the optimization problem. The curves on the left hand-side of the Figure \ref{fig: estimation_results}, are the cumulative yield curves, which represents the cumulative amount of the oil collected during each experiment. The plots on the right hand-side, represent the derivative of the cumulative yield curves. As explained in the chapter \ref{CH: Parameter_estimation}, the derivatives of the cumulative measurements are independent on the previous measurement and should be used for the parameter estimation.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautobooktabular{Figures/Results_estimation/estimation.csv} }
		\caption{Parameter estimation results rounded to fifth decimal place}
		\label{tab: Estimation_results}
	\end{table}

	The results of parameter estimation can be found in Table \ref{tab: Estimation_results} and the concentration profiles, which correspond for the optimized solution for each experiments, are presented on Figure \ref{fig: estimation_results_profiles} in Appendix \ref{CH: Profiles}. As can be noticed, the optimizer found that the partition factor should be as high as possible and moved values of $k_m$ for each experiment to unity, which was chosen as the upper bound.

	Values of the internal diffusion coefficients are distinguished for each experiment, which grows as the fluid density increase. The order of magnitude of $D_i$ obtained from the optimization is similar to values found by other researchers. \citet{Reverchon1996} performed the parameter estimation for the extraction process of sage oil from seeds and reported $D_i \approx 6e-13$.

@@

	The further spifflication of the process model can be introduced based on $C_{sat}$ results. Values of saturation concentrations are at the level of 1e5 which is an internal upper bound of the optimizer used if $+\infty$ indicates the upper bound. This suggest that solvent hasn't reach the saturation concentration and the process model can be simplified by replacing $\gamma(c_f)$ function with unity.

	As the total amount of the oil present in the system is unknown, and as such it has been obtained from the optimizer. Given the dataset and the process model, the optimizer found that the best fit is obtained if the total mass of the oil reach a lower bound equal to 80 g. The lower bound was estimated by round up of the biggest value of collected cumulative amount of the extraction product. As the same raw material was used in all experiments, the same initial amount of solute is assumed for all the cases.

	As explained above, and in Appendix \ref{CH: IC_BC}, the linear distribution of can be fully determined if the parameter $\tau$ is known. The split ratio for experiments conducted at $40 ^\circ C$ / 300 bar and $50 ^\circ C$ / 300 bar are similar to each other and close to 0.65. The experiments performed at $40 ^\circ C$ / 200 bar and $50 ^\circ C$ / 200 bar have $\tau$ equals to $0.70$ and $0.75$, respectively.

	The noise present in each dataset is quantified by parameter $\sigma$. It can be observed dataset obtained at $40 ^\circ C$ / 200 bar and $50 ^\circ C$ / 200 bar have similar value of $\sigma\approx0.36$. The visual investigation of the datasets obtained at $40 ^\circ C$ / 300 bar and $50 ^\circ C$ / 300 bar allow to explain higher values of $\sigma$ for both dataset. Both datasets have higher deviation of the data points from the simulated curve.

		\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 10.5cm 2.5cm 2.02cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P200.pdf}
			\caption{Experiment at $40^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 10.5cm 2.5cm 2.02cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P200.pdf}
			\caption{Experiment at $50^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 10.5cm 2.5cm 2.02cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P300_org.pdf}
			\caption{Experiment at $40^\circ C$ and $300$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 10.5cm 2.5cm 2.02cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P300_org.pdf}
			\caption{Experiment at $50^\circ C$ and $300$ bar}
		\end{subfigure}
		\caption{Results of parameter fitting, with estimation of the initial state}
		\label{fig: estimation_results}
	\end{figure}

	The results obtained from the parameter estimation are used to develop a generalized process model. The model generalization is performed by introducing correlations which link the estimated parameters and the operating conditions. Since the experiments were performed under varying temperature and pressure, the influence of these variables is taken into account. To simplify the problem, the correlations are written as functions of fluid density rather than temperature and pressure. Since the optimizer found that $k_m$, $m_{total}$ and $C_{sat}$ don't change with the density variations, they are assumed to be constants and not considered further.

	The other two parameters ($D_i$ and $D_e^M$) are subject to regression to obtain correlations in terms of density. Figure \ref{fig:Regression} shows the first- and second-order polynomials used to fit $D_i$ and $D_e^M$ as a function of fluid density.

	\begin{figure}[!h]
	\centering
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 11.7cm 14cm 2.3cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
		\caption{First-order polynomial regression of fitted parameters ($D_i$ and $D_e^M$) as a function of fluid density $\rho_f$}
		\label{fig:Regression_1}
	\end{subfigure}
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 11.7cm 14cm 2.3cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
		\caption{Second-order polynomial regression of fitted parameters ($D_i$ and $D_e^M$) as a function of fluid density $\rho_f$}
		\label{fig:Regression_2}
	\end{subfigure}
	\caption{Regression of $D_i$ and $D_e^M$} 
	\label{fig:Regression}
	\end{figure}

@@

	The increase in the density of the supercritical fluid can lead to an increase in the solubility of the solute and, consequently, an increase in the concentration gradient. This increased concentration gradient drives the solute to diffuse more rapidly through the matrix of the solid particles, resulting in an increased internal mass transfer coefficient. However, the effect of increased solubility on the internal mass transfer coefficient outweighs the resistance from higher density, resulting in an overall increase in the internal mass transfer coefficient with density. The same behaviour can be observed in Figure \ref{fig: estimation_results}, where plots on the right-hand side exhibit a peak. The higher the peak, the greater the increase in solute concentration in the fluid phase, which can be explained by higher solubility.

@@

	Figure \ref{fig:Regression_2} show the result of fitting a second-order polynomial in density space. It can be observed that the fit obtained from quadratic functions seems to be represent both dataset better.

	The experimental procedure and the used device can influence the values of $\tau$ and $\sigma$. Therefore, seeking a general expression for the initial state that is solely a function of operating conditions may not be sufficient to fully describe the relationship between these parameters. However, if first- and second-order regressions are applied to $\tau$ and $\sigma$, some general conclusions about the trend can be drawn. The results of the regression applied to $\tau$ and $\sigma$ are presented in Figure \ref{fig:Regression_tau_simga}.

	\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 18.8cm 8cm 2.5cm 7.6cm,clip,width=0.49\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\includegraphics[trim = 4cm 2cm 17.5cm 13.6cm,clip,width=0.49\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\caption{First order polynomial regression of fitted parameters ($\tau$ and $\sigma$) as a function of fluid density $\rho_f$}
			\label{fig:Regression_3}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 18.8cm 8cm 2.5cm 7.6cm,clip,width=0.49\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
			\includegraphics[trim = 4cm 2cm 17.5cm 13.6cm,clip,width=0.49\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
			\caption{Second order polynomial regression of fitted parameters ($\tau$ and $\sigma$) as a function of fluid density $\rho_f$}
			\label{fig:Regression_4}
		\end{subfigure}
		\caption{Regression of $\tau$ and $\sigma$} 
		\label{fig:Regression_tau_simga}
	\end{figure}

	As can be seen from Figure \ref{fig:Regression_tau_simga} , the $\tau$ seems to decrease as the fluid density increase. The standard deviation increases as the fluid density increase. As mentioned above, these results should be used with caution due to uninvestigated influence of the equipment on the extraction results.

### 2023-05-07 SFE-article/Sections/Saturation.tex
07.05

	where $k_{sat}$ is the growth rate, and it is defined as $k_{sat} = 2c_{sat}$\todo{Explain why the growth rate is defined as it is}.

@@

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 10,
					ymin = -0.5, ymax = 1.5,
					xlabel = {$c_f[kg/m^3]$},
					ylabel = {$\gamma(c_f)[-]$}]
					\addplot[
					domain = 0:10,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 2*5 .* ( x - 5 ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

>>>>

	where $k_{sat}$ is the growth rate, and it is defined as $k_{sat} = 4$. The growth rate corresponds for the shape of the $\gamma(c_f)$ function. The bigger value of the $k_{sat}$ the steeper the function is. In the limit when $ k_{sat}\rightarrow+\infty$ become a step function. If the growth rate is to high, the gradient of the $\gamma$ function become high, or goes to infinity, which might cause difficulties for a gradient-based optimizer.

@@

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 10,
					ymin = -0.5, ymax = 1.5,
					xlabel = {$c_f[kg/m^3]$},
					ylabel = {$\gamma(c_f)[-]$}]
					\addplot[
					domain = 0:10,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 4 .* ( x - 5 ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

### 2023-05-07 SFE-article/Sections/Thermo_imp.tex
07.05

	Supercritical $CO_2$'s thermodynamic properties, such as density, the local speed of sound, and specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a gas as a function of temperature and density. It can be written as a cubic function of the molar volume (of the density). Detail information about Peng-Robinson equation of state can be found in the work of \citet{Peng1976}, \citet{Elliott2011} or \citet{Pratt2001}. The P-R EOS is presented by equation \ref{EQ: PR}

@@

	The parameters $a, b, \alpha$ are parameters defined as presented in the appendix \ref{subsubsec: Equation of state}.

	The properties of $CO_2$  are presented as a function of operating conditions (temperature and pressure) in Figure \ref{fig: SFE_Properties}. At standard atmospheric pressure and temperature, $CO_2$  behaves as an ideal gas, and its compressibility factor equals unity. However, at high pressures and/or low temperatures, intermolecular forces between gas molecules become more significant, causing them to deviate from ideal behaviour. As a result, the compressibility factor can either be greater than or less than unity, depending on the magnitude of these forces. As presented in Figure \ref{fig: SFE_Properties_Compressibility}, the compressibility factor obtained from the Peng-Robinson equation of state varies strongly depending on the operating conditions. The compressibility factor can be obtained by solving the polynomial form of the P-R EOS given by \ref{EQ: PR_polynomial}.

@@

	where $A$ and $B$ are parameters as defined in the appendix \ref{subsubsec: Equation of state}. The roots of the polynomial can be found iteratively or by the Cardano formula. In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. The gas-liquid mixture is present in the two-phase region, and two roots are found. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor and its value depends on the operating conditions. The fluid's properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in the appendix \ref{subsubsec: Fluid density}.

	Figure \ref{fig: SFE_Properties_CP} show the behaviour of the heat capacity of a supercritical fluid at constant pressure ($C_p$). The details of the calculations can be found in the appendix \ref{subsubsec: Fluid heat capacity}. Contrary to the density, which varies monotonically, the specific heat shows very high levels in a narrow region. In the subcritical region, the phase transition is associated with an effective spike in the heat capacity (i.e., the latent heat). Approaching the critical point, the latent heat falls to zero, which is accompanied by a gradual rise in heat capacity in the pure phases near phase transition. At the critical point, the latent heat is zero, but the heat capacity shows a diverging singularity. Beyond the critical point, there is no divergence, but rather a smooth peak in the heat capacity; the highest point of this peak identifies the Widom line (as discussed by \citet{Simeoni2010} and \citet{Banuti2019}).

>>>>

	Supercritical $CO_2$'s thermodynamic properties, such as density, the local speed of sound, and specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a gas as a function of temperature and density. It can be written as a cubic function of the molar volume (of the density). Detail information about Peng-Robinson equation of state can be found in the work of \citet{Peng1976}, \citet{Elliott2011} or \citet{Pratt2001}. The P-R EOS is presented by Equation \ref{EQ: PR}

@@

	The parameters $a, b, \alpha$ are parameters defined as presented in the Appendix \ref{subsubsec: Equation of state}.

	The properties of $CO_2$  are presented as a function of operating conditions (temperature and pressure) in Figure \ref{fig: SFE_Properties}. At standard atmospheric pressure and temperature, $CO_2$  behaves as an ideal gas, and its compressibility factor equals unity. However, at high pressures and/or low temperatures, intermolecular forces between gas molecules become more significant, causing them to deviate from ideal behaviour. As a result, the compressibility factor can either be greater than or less than unity, depending on the magnitude of these forces. As presented in Figure \ref{fig: SFE_Properties_Compressibility}, the compressibility factor obtained from the Peng-Robinson equation of state varies strongly depending on the operating conditions. The compressibility factor can be obtained by solving the polynomial form of the P-R EOS given by Equation \ref{EQ: PR_polynomial}.

@@

	where $A$ and $B$ are parameters as defined in the Appendix \ref{subsubsec: Equation of state}. The roots of the polynomial can be found iteratively or by the Cardano formula. In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. The gas-liquid mixture is present in the two-phase region, and two roots are found. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor and its value depends on the operating conditions. The fluid's properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in the Appendix \ref{subsubsec: Fluid density}.

	Figure \ref{fig: SFE_Properties_CP} show the behaviour of the heat capacity of a supercritical fluid at constant pressure ($C_p$). The details of the calculations can be found in the Appendix \ref{subsubsec: Fluid heat capacity}. Contrary to the density, which varies monotonically, the specific heat shows very high levels in a narrow region. In the subcritical region, the phase transition is associated with an effective spike in the heat capacity (i.e., the latent heat). Approaching the critical point, the latent heat falls to zero, which is accompanied by a gradual rise in heat capacity in the pure phases near phase transition. At the critical point, the latent heat is zero, but the heat capacity shows a diverging singularity. Beyond the critical point, there is no divergence, but rather a smooth peak in the heat capacity; the highest point of this peak identifies the Widom line (as discussed by \citet{Simeoni2010} and \citet{Banuti2019}).
